% =====================================================================
% DROP-IN SECTION for ILAC-WAN_BTP_Proposal_v2.tex
% Insert with \input{freshness_section} right after the Direction 2
% subsection (line ~486), or as a standalone \section before
% "Comparative Assessment". It uses the proposal's own macros
% (\Dir, \Ltag, limitbox, ideabox, goldbox) and color palette, so it
% compiles unchanged on Overleaf. New \cite keys are in freshness_refs.bib.
% Also recommended: edit the Executive Summary / Introduction to state the
% spine sentence (see freshness_intro_patch.tex).
% =====================================================================

\section{Implemented Contribution: The Moving-Target Freshness Spine}
\label{sec:freshness}

The professor's decisive feedback was that the work needed a single, clear
objective. We provide one and implement it: \emph{energy-optimal semantic
knowledge aggregation for swarms observing \textbf{moving} targets}, with target
motion entering through \textbf{information freshness}. This makes
\Dir{1} (fidelity-aware compression) and \Dir{2} (predictive topology) two
mechanisms of one story rather than two independent tweaks, and it activates the
exact term the audit flagged as missing in \Ltag{L8}: \emph{``no fidelity,
freshness, or value of information.''}

\subsection{Why motion must enter through value, not mobility (a verified negative result)}

The intuitive integration --- re-anchor each patrol centre $\mathbf c_i$ on the
(predicted) target and let agents physically chase --- is \emph{inert} in this
framework, and we show it empirically in our own faithful re-implementation.

\begin{limitbox}{Mobility-driven tracking does nothing here}
At the paper's patrol radius $R\in[80,100]$\,m, a target moving $3$--$7$\,m/s
travels only $30$--$41$\,m per round, a fraction of $R$, so the coverage
constraint $\lVert\mathbf p_i-\mathbf c_i\rVert\le R$ never binds: target class
is invisible and the apparent energy differences are non-monotone noise (std
$\gg$ signal). Decoupling a tight sensing radius $r_{\mathrm{track}}=8$\,m makes
the \emph{geometric} chase demand explode to $40$\,m/round, yet realized energy
stays identical to the static case ($0.597$ vs.\ $0.601$\,J), because the inner
solver only \emph{targets} coverage during its optional motion step and the
matcher routes around it. Mobility is a near-discontinuous radio-tuning term, not
a tracking cost.
\end{limitbox}

This is consistent with the proposal's own ranking: the mobility/sensing route is
\Dir{4} (ISAC co-design), demoted to rank~4 precisely because Cram\'er--Rao-bound
trajectory design is crowded, and the 2024--2026 ISAC-tracking literature
(EKF\,+\,posterior-CRB trajectory optimization) confirms that lane is saturated
\cite{isacswarm}. We therefore couple motion to the framework's \emph{own}
objects --- the compression ratio $\eta$ and the aggregation tree.

\subsection{The freshness model}

Each source $s$ (data collected by agent $s$) observes a target with
\emph{agility} $a_s$, its mean speed over the horizon. The receiver's estimation
error of a moving target is a non-decreasing function of the age of the last
update --- the standard Age-of-Information / remote-estimation result
\cite{sun2017,aoii2020,aoi}. We price that staleness as a freshness distortion
that grows with how long a source's data waits before delivery,
\begin{equation}
D^{\mathrm{fresh}}_s \;=\; a_F\, a_s\,(\tau_s)^p,\qquad p\ge 1,
\label{eq:dfresh}
\end{equation}
where $\tau_s = h_s\,T$ is the source's age in aggregation hops $h_s$ (each
fuse--compress--relay stage is one processing-age step --- the multi-hop /
version-AoI model \cite{aosi2025}) and $p{=}1$ is the linear AoI form. Added to
the path-multiplicative compression distortion of \Dir{1} (audit \Ltag{L14}), the
assembled root report carries
\begin{equation}
D_{\mathrm{root}} \;=\; \underbrace{a_D\textstyle\sum_s w_s\!\!\sum_{\text{hops}}\!\ln(1/\eta)}_{\text{compression (Dir.\,1)}}
\;+\; \underbrace{\textstyle\sum_s w_s\, D^{\mathrm{fresh}}_s}_{\text{freshness (new)}},\qquad
w_s=\tfrac{\text{bits}_s}{\sum_{s'}\text{bits}_{s'}}.
\label{eq:droot}
\end{equation}
Equation~\eqref{eq:droot} couples motion to two controls: the compression ratio
$\eta$ (per pair) and the per-source hop count $h_s$ (the topology).

\subsection{Two coupling points, both verified}

\begin{ideabox}{\Dir{1} --- compression bends to motion}
Through the endogenous per-hop $\eta$-floor (our replacement for the exogenous
$\eta^{\mathrm{req}}$, audit \Ltag{L23}), a faster target consumes more of the
fidelity budget $D_{\max}$ via \eqref{eq:dfresh}, so the floor \emph{rises} and
the agent must compress \emph{less}. Exactly: as target speed goes $0\!\to\!9$\,m/s
the optimal $\eta$ rises $0.549\!\to\!0.719$ and energy rises $+9\%$, monotone and
reproducible (not noise), and robust across $30$ random pairs. This strengthens
the headline \Dir{1} result: the paper's ``always compress maximally'' fails by an
amount set by \emph{target motion}.
\end{ideabox}

\begin{ideabox}{\Dir{2} --- a fresher report, for free}
With $D_{\mathrm{root}}$ depending on $h_s$, the matching weight gains the
marginal staleness a source accrues when its holder transmits:
$w(i\!\to\!j)=E(i\!\to\!j)+\lambda_F\,a_F\,T\!\sum_{s\in i}\text{bits}_s\,a_s$.
The same exact Blossom matcher then gives fast-decaying sources shorter paths.
Over $20$ missions this cuts root staleness by \textbf{40\%} at \textbf{no energy
cost} ($-1.5\%$), reduces \emph{exactly} to the paper at $\lambda_F{=}0$
($20/20$ identical), and its advantage grows with agility spread
($21\%\!\to\!48\%$) --- the goal-oriented / value-of-information regime
\cite{goaloriented2025}.
\end{ideabox}

\paragraph{Prediction (effective decay).} The agility in \eqref{eq:dfresh} is the
predictor's residual error rate, so a better predictor lowers it. A
constant-velocity Kalman filter and a CV/CT interacting-multiple-model (IMM)
predictor --- the established maneuvering-target tools --- dominate the naive
finite-difference predictor on every target class; predicting ahead pays once
motion exceeds sensor noise and yields a ${\sim}23\%$ fresher report. Honest
scope: for near-static targets, not predicting is best, so the freshness-optimal
agent simply selects the lowest-residual predictor.

\subsection{Metaheuristic matchers, used correctly (mentor's request)}

The per-round pairing is solved \emph{exactly} by Blossom in polynomial time, so
nature-inspired metaheuristics cannot beat it at this scale --- they are included
only as honest comparison baselines. We implement ant colony, artificial bee
colony, cuckoo search, and Egyptian vulture over the pairing objective. They
approach but never beat exact matching (a $0.1$--$0.3\%$ gap that grows with $N$
while Blossom stays optimal and fast, $12$\,ms at $N{=}30$), and at mission level
they cluster at the myopic per-round optimum ($\approx 13$--$15\%$ above the
brute-force oracle) while only the learned \emph{lookahead} policy closes the gap
($3.6\%$). The contribution is the learning, not the metaheuristic --- which is
exactly the defensible conclusion.

\subsection{Verification}

\begin{goldbox}{Every claim is an automated check}
The moving-target spine adds $11$ checks (\texttt{run\_freshness.py}) and the
metaheuristic baselines add $5$ (\texttt{run\_metaheuristics.py}); with the
existing suites the project totals \textbf{58 automated checks, $0$ failures},
each regenerating from code in minutes. A polished, animated web demo replays
these real computed traces for the viva and reviewers.
\end{goldbox}

\noindent\textbf{Headline numbers (all measured).} $\eta$ rises $0.549\!\to\!0.719$
and energy $+9\%$ with target speed; value-prioritized topology is $40\%$ fresher
at no energy cost and reduces to the paper at $\lambda_F{=}0$; the freshness
advantage scales $21\%\!\to\!48\%$ with agility spread; learned topology gap
$3.6\%$ vs.\ paper $14.7\%$ vs.\ random $29.1\%$; Kalman/IMM dominate naive
prediction with a ${\sim}23\%$ fresher report.
